The Implicit Calculus: A New Foundation for Generic Programming

### 2017-01-30 src/Translation.tex

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
\paragraph{Resolution}
The more interesting part of the translation happens when resolving queries.
Queries are translated by rule $\TyQuery$ using the auxiliary resolution
judgment $\vturns$:
\begin{equation*}
\tenv \vturns \rulet~\gbox{\leadsto E}
\end{equation*}
which is shown, in deterministic form, in Figure~\ref{fig:resolution2}.
The translation builds a System F term as evidence for the resolution.
The mechanism that builds evidence dualizes the process of peeling off
abstractions and universal 
quantifiers. The rule \mylabel{R-IAbs}~wraps a lambda binder with a fresh variable 
$x$ around a System F expression $E$, which is generated from the resolution 
for the head of the rule ($\rulet_2$). The rule \mylabel{R-TAbs}~wraps a type lambda binder 
around the System F expression resulting from the resolution of $\rulet$.

TODO

For simple types \mylabel{R-Simp}, evidence $E_1$ is retrieved from matching 
$\type$ and then used in the resolution process of the simple type 
$\type$ with rule $\turns_\downarrow$:

\begin{center}
$\bar{\alpha};\env; \rulet~\gbox{\leadsto E_1} \turns_\downarrow \type~\gbox{\leadsto E_2}$
\end{center}

The rules \mylabel{I-IAbs}~and \mylabel{I-TAbs}~are the most interesting in the translation. 
In rule \mylabel{I-IAbs}~we need evidence for $\rulet_2$ to build evidence for $\type$, 
but all we have is evidence $E_1$ for $\rulet_1 \iarrow \rulet_2$. However, we can 
construct evidence for $\rulet_2$, by generating evidence $E_2$ for $\rulet_1$
and then simply applying $E_1$ to $E_2$. Rule \mylabel{I-TAbs}~is similar: we need 
evidence for $\rulet$, but all we have is evidence $E_1$ for $\forall \alpha. \rulet$. 
To create evidence for $\rulet$ we can substitute $\alpha$ by some type 
$\rulet'$ in $\rulet$ and generate a term which is the type application of 
$E_1$ to $|\rulet'|$.

Finally, matching ($\elookup{\env}{\type}$) also needs to be
extended with evidence generation, but this extension is
straightforward.

\paragraph{Meta-Theory} The type-directed translation of $\ourlang$ to System F exhibits a number
of desirable properties.

>>>>

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
\paragraph{Resolution}
The more interesting part of the translation happens when resolving queries.
Queries are translated by rule $\TyQuery$ using the auxiliary resolution
judgment $\ivturns$:
\begin{equation*}
\tenv \ivturns \rulet~\gbox{\leadsto E}
\end{equation*}
which is shown, in deterministic form, in Figure~\ref{fig:resolution2}.  The
translation builds a System F term as evidence for the resolution.

The mechanism that builds evidence dualizes the process of peeling off
abstractions and universal quantifiers: Rule \mylabel{R-IAbs}~wraps a lambda
binder with a fresh variable $x$ around a System F expression $E$, which is
generated from the resolution for the head of the rule ($\rulet_2$). Similarly,
rule \mylabel{R-TAbs}~wraps a type lambda binder around the System F expression
resulting from the resolution of $\rulet$.

For simple types rule $\type$ rule \mylabel{R-Simp} delegates the work of
building evidence, when a matching rule $\rulet$ type is found in the
environment, to rule \mylabel{L-RuleMatch}. The evidence consists of two parts:
$E$ is the evidence of matching $\type$ against $\rulet$. This match contains
placeholders $\bar{x}$ for the contexts whose resolution is postponed by rule
\mylabel{M-IAbs}. It falls to rule \mylabel{L-RuleMatch} to perform these
postponed resolutions, obtain their evidence $\bar{E}$ and fill in the
placeholders.

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
\paragraph{Meta-Theory} The type-directed translation of $\ourlang$ to System F exhibits a number
of desirable properties.
